Convergence rate of MAP estimates \\ for the exponential family

### 2021-04-15 main.tex
Apply the log lower bound to the MLE

\begin{example}
	\paragraph{MAP upper bound.}
	Wolfram finds no closed form for $\expect{\log\frac{\MAPm}{\mu^*}}$ but it does for $\expect{\frac{\mu^*}{\MAPm}} = \expect{\frac{\MAPt}{\natp^*}}$. 
	To start, let us plug in the definition of $\MAPm$
	\begin{align}
		\E[\MAPt] &= -\frac{n+n_0}{2}\expect{\inv{n_0 \mu_0 + \sum_i X_i^2}}  \\
		&= -\frac{n+n_0}{2 \mu^*}\expect{\inv{n_0 \frac{\mu_0}{\mu^*}+ \sum_i \frac{X_i^2}{ \mu^*} }} \\
		&= -\frac{n+n_0}{2 \mu^*}\expect{\inv{a + \chi^2(n)}}
	\end{align}
	where $\sum_i \frac{X_i^2}{ \mu^*}$ is a chi-square random variable of degree $n$ and 
	\begin{align}
			a=n_0 \frac{\mu_0}{\mu^*} \; .
	\end{align}
	This expectations is not simple to solve, but using wolfram, we get  the formula :
	\begin{align}
		\expect{\inv{a + \chi^2(n)}}  = \half e^{\frac{a}{2}} E_{\half[n]}(\half[a])
		\label{eq:wolfram_input}
	\end{align}
	where the \href{https://dlmf.nist.gov/8.19}{generalized exponential integral function} is defined as 
	\begin{align}
		E_k(z) = \int_1^\infty \frac{e^{-z t} }{t^k} dt \; .
	\end{align}
	Now our goal is to bound this function with simpler functions.
	Fortunately, mathematicians have been working on these integrals for decades.
	When $n=2$, we have the simple exponential integral function which \href{https://dlmf.nist.gov/6.8#E1}{verifies the bound}
	\begin{align}
		\half \log(1 + \frac{2}{x}) 
		\leq e^x E_1(x) 
		\leq \log(1+\inv{x}) \leq \inv{x} \; .
	\end{align}
	For larger $n$, we have \href{https://dlmf.nist.gov/8.19#E21}{the general bound}
	\begin{align}
		\inv{x+n}
		\leq e^x E_k(x) 
		\leq \inv{x + k - 1}
	\end{align}
	so that 
	\begin{align}
		\inv{a+n}
		\leq \expect{\inv{a + \chi^2(n)}}
		\leq \inv{a + n - 2}
	\end{align}
	so that
	\begin{align}
		\frac{n +n_0}{2 \mu^* (a + n)}	
		&\leq - \E[\MAPt]
		\leq \frac{n +n_0}{2 \mu^* (a + n - 2)}	\\
		\iff \inv{2 \mu_n}
		&\leq - \E[\MAPt]
		\leq \inv{2 \mu_n - \frac{4 \mu^*}{n + n_0}}	\\
		\iff \frac{\mu^*}{\mu_n}
		&\leq \expect{\frac{\mu^*}{\MAPm}} 
		\leq \frac{\mu^*}{\mu_n - \frac{4 \mu^*}{n + n_0}}  
		\label{eq:theta_expectation_bound}
	\end{align}
	because $\frac{\mu^* (a + n)}{n +n_0} = \mu_n$.
Now using the bound \eqref{eq:log_bound} on the logarithm, we get
\begin{align}
	2\expect{\bregmanconj(\mu^*; \MAPm)} 
	&\leq \expect{\frac{\mu^*}{\MAPm}}  + \expect{\frac{\MAPm}{\mu^*}} - 1 \\
	&\leq \frac{\mu^*}{\mu_n - \frac{4 \mu^*}{n + n_0}}   + \frac{\mu_n}{\mu^*} - 1 \\
	&= \frac{n_0 + n}{a +n - 2} + \frac{a+n}{n_0 + n} - 2 \\
	&= \frac{n_0 - a + 2}{a +n - 2} + \frac{a - n_0}{n_0 + n} \\
	&= \frac{2}{a+n-2} + \frac{(n_0 - a)^2}{(a+n-2)(n_0 + n)}
\end{align}
where we remind you that $a=n_0\frac{\mu_0}{\mu^*}$. This inequality highlights a clear variance-bias decomposition, to compare with the MLE.
\end{example}

>>>>

\begin{example}
	\paragraph{MAP upper bound.}
	Wolfram finds no closed form for $\expect{\log\frac{\MAPm}{\mu^*}}$ but it does for $\expect{\frac{\mu^*}{\MAPm}} = \expect{\frac{\MAPt}{\natp^*}}$. 
	To start, let us plug in the definition of $\MAPm$
	\begin{align}
		\E[\MAPt] &= -\frac{n+n_0}{2}\expect{\inv{n_0 \mu_0 + \sum_i X_i^2}}  \\
		&= -\frac{n+n_0}{2 \mu^*}\expect{\inv{n_0 \frac{\mu_0}{\mu^*}+ \sum_i \frac{X_i^2}{ \mu^*} }} \\
		&= -\frac{n+n_0}{2 \mu^*}\expect{\inv{a + \chi^2(n)}}
	\end{align}
	where $\sum_i \frac{X_i^2}{ \mu^*}$ is a chi-square random variable of degree $n$ and 
	\begin{align}
			a=n_0 \frac{\mu_0}{\mu^*} \; .
	\end{align}
	This expectations is not simple to solve, but using wolfram, we get  the formula :
	\begin{align}
		\expect{\inv{a + \chi^2(n)}}  = \half e^{\frac{a}{2}} E_{\half[n]}(\half[a])
		\label{eq:wolfram_input}
	\end{align}
	where the \href{https://dlmf.nist.gov/8.19}{generalized exponential integral function} is defined as 
	\begin{align}
		E_k(z) = \int_1^\infty \frac{e^{-z t} }{t^k} dt \; .
	\end{align}
	Now our goal is to bound this function with simpler functions.
	Fortunately, mathematicians have been working on these integrals for decades.
	When $n=2$, we have the simple exponential integral function which \href{https://dlmf.nist.gov/6.8#E1}{verifies the bound}
	\begin{align}
		\half \log(1 + \frac{2}{x}) 
		\leq e^x E_1(x) 
		\leq \log(1+\inv{x}) \leq \inv{x} \; .
	\end{align}
	For larger $n$, we have \href{https://dlmf.nist.gov/8.19#E21}{the general bound}
	\begin{align}
		\inv{x+n}
		\leq e^x E_k(x) 
		\leq \inv{x + k - 1}
	\end{align}
	so that 
	\begin{align}
		\inv{a+n}
		\leq \expect{\inv{a + \chi^2(n)}}
		\leq \inv{a + n - 2}
	\end{align}
	so that
	\begin{align}
		\frac{n +n_0}{2 \mu^* (a + n)}	
		&\leq - \E[\MAPt]
		\leq \frac{n +n_0}{2 \mu^* (a + n - 2)}	\\
		\iff \inv{2 \mu_n}
		&\leq - \E[\MAPt]
		\leq \inv{2 \mu_n - \frac{4 \mu^*}{n + n_0}}	\\
		\iff \frac{\mu^*}{\mu_n}
		&\leq \expect{\frac{\mu^*}{\MAPm}} 
		\leq \frac{\mu^*}{\mu_n - \frac{4 \mu^*}{n + n_0}}  
		\label{eq:theta_expectation_bound}
	\end{align}
	because $\frac{\mu^* (a + n)}{n +n_0} = \mu_n$.
Now using the bound \eqref{eq:log_bound} on the logarithm, we get
\begin{align}
	2\expect{\bregmanconj(\mu^*; \MAPm)} 
	&\leq \expect{\frac{\mu^*}{\MAPm}}  + \expect{\frac{\MAPm}{\mu^*}} - 1 \\
	&\leq \frac{\mu^*}{\mu_n - \frac{4 \mu^*}{n + n_0}}   + \frac{\mu_n}{\mu^*} - 1 \\
	&= \frac{n_0 + n}{a +n - 2} + \frac{a+n}{n_0 + n} - 2 \\
	&= \frac{n_0 - a + 2}{a +n - 2} + \frac{a - n_0}{n_0 + n} \\
	&= \frac{2}{a+n-2} + \frac{(n_0 - a)^2}{(a+n-2)(n_0 + n)} \\
	\implies \expect{\bregmanconj(\mu^*; \MAPm)} 
	&\leq \frac{1}{a+n-2} + \frac{(n_0 - a)^2}{2(a+n-2)(n_0 + n)} 
\end{align}
where recall that $a=n_0\frac{\mu_0}{\mu^*}$. 
This inequality highlights a clear variance-bias decomposition, to compare with the MLE. Unfortunately it is not tight asymptotically as the inequality \eqref{eq:log_bound} is far from tight. We are basically losing a factor 2 compared to $1/2n$.

\paragraph{MLE Loose Upper Bound.}
If we apply \eqref{eq:log_bound} to the MLE instead, we get the upper bound
\begin{align}
	\expect{\bregmanconj(\mu^*; \MAPm^\text{MLE})}
	&\leq  \half \left( \expect{\frac{n}{\chi^2(n)}} + \expect{\frac{\chi^2(n)}{n}} - 2 \right ) \\
	&= \half \left( \frac{n}{n-2} + 1 - 2 \right)  
	= \frac{1}{n-2}
\end{align}
which is exactly $1/2n$ larger than \eqref{eq:tight_bounds} which we got thanks to an exact computation of the expectation of the log term, and a tight upper bound on the digamma function.
\end{example}
